A Survey of Near-Term Exoskeleton Technology\\ (Draft 3.0)

### 2015-09-28 exos/exosText_bodyExt.tex

The body extender is one of the few existing full-body exoskeletons.  The Body Extender's main objective is to increase the forceful interaction capabilities, specifically handling heavy loads, of an operator in difficult and unknown environments.  The overall suit is composed of four robotic limbs with kinematics designed to be anthropomorphically similar.  The suit contains twenty-two independently actuated degrees of freedom.  Note that the Body Extender is the only fully actuated suit included in this report.  The fully actuated design of the suit was chosen based on the fact that handling, or rather manipulating heavy objects requires forces and torques that far exceed human capabilities.  Note also that having many actuated degrees of freedom means that fully actuated suits will be very heavy, and thus will require specialized controllers.

The main design objectives for the Body Extender included  requiring that operators be able to use the hardware with a minimal amount of training.  Thus, the system should not require any substantial modification the a human's motor habits.  This objective includes the fact that the mass distribution of the suit, of special consideration in these heavy systems, be similar to that of an unloaded operator.  Additionally, the resistive forces between the operator and suit should always be minimized, even when the suit is not loaded externally.

A picture of the Body Extender exoskeleton and its five main components, four limbs and a torso, are shown in Figure \ref{fig:bodyExt}.
\begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/bodyExt/bodyExt}
  \caption{}
  \vspace{-0.2in}
 \label{fig:bodyExt}   
 \end{figure}

@@

 Each of the actuators in the Body Extender suit, execpt the forearm pronosupination are dirven by a linear actuator which is composed of an electrical motor, incremental encoder, ball screw, and angular contact ball bearings.  The motors are frameless DC torque controlled motors.  The {\bf peak torque of the motors are 7 Nm, with a maximum continuous torque of 6 Nm}.  Each motor weighs 1.4 kg.  The {\bf linear actuator assembly is able to supply 8000 N with a total weight of 2.4 kg}.

 For joints that have a small range of motion and correspondingly low torque requirements, the linear actuators are used in straightforward lever mechanism configurations.  For degrees of free which require relatively large ranges of motion as well as torque, a motion conversion system is used to transform the linear actuator motion into rotational motion.  The motion conversion unit consisted of a pantograph, two metallic tendons, and an output pulley connected to the output link.  A CAD drawing of the motion conversion unit is shown in Figure \ref{fig:motionConv}. 
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/bodyExt/motionConv}
  \caption{}
%  \vspace{-0.2in}
 \label{fig:motionConv}   
 \end{figure}
 There are two configurations of the motion conversion unit.  The first has a {\bf maximum continuous output of 500 Nm and the second a maximum continuous output of 270 Nm}.  The two units weigh 6 kg and 5 kg respectively, and have {\bf mechanical efficiency of about 85\%}.  {\bf The speed of the actuators are about 60 deg/s}.

@@

 The Body Extender exoskeleton consists of twenty-two independently actuated degrees of freedom.  The suit has two arms, two legs,and a torso, or central backpack unit.

 Each of the two arms on the system has a servo-amplified degree of freedom for grasping heavy objects.  The grasping force of various objects is achieved using a force sensing trigger mechanism; the grasping force is proportional to the force applied to the trigger.

 The suit contains twenty-two incremental encoders which measure angular displacement of each of the electric motors.  There are also {\bf five six-axis force/torque sensors mounted at the five connecting points between the user and the device; two at the hands, two at the feet, and one on the trunk.}  These sensors directly measure the forces and torques between the user and the suit.  There are additionally two single axis force sensors integrated into the grippers that measure the force applied by the operator on the ``gripper triggers," as well as {\bf a three axis accelerometer integrated into the backpack unit which estimates the trunk orientation}.  Lastly, accelerometers are distributed throughout the system to monitor link dynamics, serving the roll of supporting overall equilibrium and internal stability during manipulation tasks.

{\bf The control software is distributed on one central processing unit as well as sixteen local units located throughout the system}.  The communication between processing units is handled using a high-speed field bus.

{\bf The system is powered using thirteen batteries in series, which provide a nominal voltage of 78 V}.  When the system is run in stand alone mode, the batteries can provide eight hours of continuous work.

@@

 The body extender control algorithm centers around a {\bf feedback linearization policy} that depends on measuring the force between the operator and suit, knowing the contact state of the suit and environment, decoupling the control of each of the limbs for each other, and having each limb be fully actuated.  The dynamics for each of the Body Extender limbs are written in standard form as  
\begin{equation}
B_i(q) \ddot{q} + C_i(q,\dot{q})\dot{q} + G_i(q) = \tau + J^T_{iu}F_s + J^T_{ia}F_a,
\end{equation}
where the $i = \{ i,2,3,4\}$ where each integer corresponds to a different limb, $\tau$ represents the actuator torques, $F_s$ the measured forces between the operator and suit, and $F_a$ the unmeasured contact forces on the suit.  The $B_i$, $C_i$, and $G_i$ terms represent the limb inertia, Coriolis and centrifugal, and gravitational torques.

The authors in \cite{} use a decoupling strategy in their controllers which enables the individual limbs in the Body Extender to be somewhat decoupled from each other.  They claim that they accomplish this using the following model:
\begin{equation}
B_i(q) \ddot{q} + C_i(q,\dot{q})\dot{q} + G_{iT}(q) = \tau + J^T_{iu}F_s + J^T_{ia}F_a - B_{iT}{q}\dot{V}_T - C_{iT}(q,\dot{q})V_T,
\end{equation}
where, $V_T$ is the trunk velocity, $G_{iT}$ the gravitational effect of the trunk on the limb, $B_{iT}$ the inertia of the trunk acting on the limb, and $C_{iT}$ the trunk Coriolis and centrifugal term.  Note that $V_T$ and $\dot{V}_T$ are directly measured through the accelerometer and gyroscope in the trunk.

The basis of the feedback linearizing controller in \cite{} operates under two fundamental assumptions, and only appears to relate directly to the case of a single one of the manipulator arms.  The first assumption of the controller is that the mass and inertial properties of an object being manipulated are estimated online.  The second is that the forces measured between the operator and suit can be modeled as
\begin{equation}
F_s = M_p \dot{V}_e + G_p,
\label{eq:forceEq}
\end{equation}
where $M_p$ is an effective mass of the end effector, $\dot{V}_e$ is the acceleration of the end effector, and $G_p$ is an effective gravitational term.  Note that the gravitational term is zero if the arm is not holding an external load.  The importance of these two assumptions will be more clear after the analytical forms of the feedback linearizing controllers are introduced.

The dynamics of the arm when not manipulating a load are
\begin{equation}
B(q) \ddot{q} + C(q,\dot{q})\dot{q} + G(q) = \tau. \notag
\end{equation} 
The feedback linearizing controller for this case has the form
\begin{equation}
\tau = B(q)(\ddot{q}_d + K_v \dot{\tilde{q}} + K_p \tilde{q}) + C(q,\dot{q})\dot{q} + G(q), \notag
\end{equation}
where $\tilde{q} = q_d - q$.  The details are unclear in \cite{}, but it appears that the authors either propose using predefined trajectories to define $\ddot{q}_d$, $\dot{q}_d$, and $q_d$, or use the assumption in \eqref{eq:forceEq} to determine these values.  
%\begin{equation}
%F_s = M_p \dot{V}_e
%\label{eq:assump1}
%\end{equation}
Equation \ref{eq:forceEq} relates the measured force to the derivative of the end effector velocity $\dot{V}_e$.  This end effector velocity can then be related to the desired joint velocities using the analytical jacobian transformation \[ \ddot{q}_d = J(q)^{-1}(\dot{V}_e - \dot{J}(q,\dot{q})\dot{q}).\]  The values for $\dot{q}_d$ and $q_d$ can then be found through integration.  It is not clear the how authors explicitly decouple the human dynamics from the robot in this case.

In the case where the exoskeleton arm is manipulating a load, the feedback linearizing controller has the augmented form
\begin{equation}
\tau = B(q)(\ddot{q}_d + K_v \dot{\tilde{q}} + K_p \tilde{q}) + C(q,\dot{q})\dot{q} + G(q) -J^T(q) F_s -J^T(q)F_a. \notag
\end{equation}
The main difference in this case is that the controller here explicitly compensates for the measured human-robot interaction force $F_s$, and the controller includes the un-sensed forces $F_a$ which arise from the object being manipulated.  The object interaction forces thus need to be estimated, which was the first fundamental assumption on which the control scheme for the Body Extender relied.  Because the point at which the manipulator arm and object are connected is assumed to be well known, and the arm is instrumented with inertial sensors, estimating the interaction forces of the manipulated object is be assumed to be equivalent to estimating its inertial properties.  As outlined in \cite{}, this is accomplished using a standard regressor which uses the input and output of the feedback linearizing controller as the basis for the regression.  The details of this process are left to \cite{}.

@@

The Body Extender suit is one of the only full-body exoskeletons available today, and as such is a somewhat remarkable system. In terms of how the Body Extender system is designed and how it is controlled, the suit may be limited. Clearly, for the task of handling heavy objects a fully actuated suit may be the best option, but generally speaking this seems almost like an over engineered solution.  The Body Extender is a heavy, overly clunky suit.  This suit may be alright for quasi-static operations, but it is unlikely that it will be useful for more than pick and place type operations.  Again, seeing as this is basically what it was designed for, that seems alright.

@@

In terms of the control of the overall suit, it was difficult to find any control papers about the Body Extender that covered anything other than upper-body single-arm control of a manipulated object.  The decoupling strategy discussed above, decoupling the individual limbs from each other, seems somewhat feasible, but is not the best way to approach coordinated limb control.  It was difficult to find literature documenting the specifics of lower-limb control for the Body Extender.  Presumably a similar strategy to that discussed here, with the addition of gravity compensation for the suit, is used.

>>>>

The Body Extender is one of the few existing full-body exoskeletons.  The Body Extender's main objective is to increase the forceful interaction capabilities, specifically the heavy load handling capabilities, of an operator in difficult and unknown environments.  The overall suit is composed of four robotic limbs with kinematics designed to be anthropomorphically similar.  The suit contains twenty-two independently actuated degrees of freedom.  Note that the Body Extender is the only fully actuated suit included in this report.  The fully actuated design of the suit was chosen based on the fact that handling, or rather manipulating heavy objects requires forces and torques that far exceed human capabilities.  Note also that having many actuated degrees of freedom means that fully actuated suits will be very heavy, and thus will require specialized controllers.

As a primary objective, the Body Extender aims to allow operators to use the hardware with minimal training.  Thus, the system is designed so as not to require substantial modifications to human motor habits.  The effective mass distribution of the suit, of special consideration in these heavy systems, is designed to be similar to that of an unloaded operator.  Like most other strength augmentation exoskeletons, the Body Extender is intended to minimize resistive forces between the operator and suit, even when the suit is not loaded externally.

A picture of the Body Extender exoskeleton and its five main components, four limbs and a torso, is shown in Figure \ref{fig:bodyExt}.
\begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/bodyExt/bodyExt}
  \caption{}
  \vspace{-0.2in}
 \label{fig:bodyExt}   
 \end{figure}

@@

 Except for forearm pronosupination, each of the actuators in the Body Extender suit are driven by a linear actuator that is composed of an electrical motor, incremental encoder, ball screw, and angular contact ball bearings.  The motors are frameless DC torque controlled motors.  The {\bf peak torque of the motors is 7 Nm, with a maximum continuous torque of 6 Nm}.  Each motor weighs 1.4 kg.  The {\bf linear actuator assembly is able to supply 8000 N with a total weight of 2.4 kg}.

 Joints that have a small range of motion and low torque requirements use linear actuators in straightforward lever mechanism configurations.  For degrees of freedom which require relatively large ranges of motion as well as torque, a motion conversion system transforms the linear actuator motion into rotational motion.  The motion conversion unit consists of a pantograph, two metallic tendons, and an output pulley connected to the output link.  A CAD drawing of the motion conversion unit is shown in Figure \ref{fig:motionConv}. 
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/bodyExt/motionConv}
  \caption{}
%  \vspace{-0.2in}
 \label{fig:motionConv}   
 \end{figure}
 There are two configurations of the motion conversion unit.  The first has a {\bf maximum continuous output of 500 Nm and the second a maximum continuous output of 270 Nm}.  The two units weigh 6 kg and 5 kg respectively, and have {\bf mechanical efficiency of about 85\%}.  {\bf The speed of the actuators is approximately 60 deg/s}.

@@

 The Body Extender's twenty-two actuated degrees of freedom include two arms, two legs, and a torso or central backpack unit.   The arms have a servo-amplified degree of freedom for grasping heavy objects.  Grasping forces are achieved using a force sensing trigger mechanism; the force is proportional to the force applied to the trigger.

 The suit contains twenty-two incremental encoders that measure angular displacement of each of the electric motors.  There are also {\bf five six-axis force/torque sensors mounted at the five connecting points between the user and the device; two at the hands, two at the feet, and one on the trunk.}  These sensors directly measure the forces and torques between the user and the suit.  There are additionally two single axis force sensors integrated into the grippers that measure the force applied by the operator on the ``gripper triggers," as well as {\bf a three axis accelerometer integrated into the backpack unit which estimates the trunk orientation}.  Lastly, \textbf{accelerometers are distributed throughout the system} to monitor link dynamics, supporting overall equilibrium and internal stability during manipulation tasks.

{\bf The control software is distributed between a central processing unit and sixteen local units located throughout the system}.  The communication between processing units is handled using a high-speed field bus.

{\bf The system is powered using thirteen batteries in series, which provide a nominal voltage of 78 V}.  When the system is run in stand-alone mode, the batteries can provide eight hours of continuous work.

@@

 The body extender control algorithm centers around a {\bf feedback linearization policy} that depends on measuring the force between the operator and suit, knowing the contact state of the suit and environment, decoupling the control of each of the limbs for each other, and having each limb be fully actuated.  The dynamics for each of the Body Extender limbs are written in standard form as  
\begin{equation}
\mB_i(q) \ddot{q} + \mC_i(q,\dot{q})\dot{q} + \mG_i(q) = {\bf \tau} + \mJ^T_{iu}{\bf F_s} + \mJ^T_{ia}{\bf F_a},
\end{equation}
where $i$ is an integer corresponding to a different limb, ${\bf \tau}$ represents the actuator torques, ${\bf F_s}$ the measured forces between the operator and suit, and ${\bf F_a}$ the unmeasured contact forces on the suit.  The $\mB_i$, $\mC_i$, and $\mG_i$ terms represent the limb inertia, Coriolis and centrifugal, and gravitational torques.

The authors in \cite{} use a decoupling strategy in their controllers which enables the individual limbs in the Body Extender to be somewhat decoupled from each other.  They claim that they accomplish this using the following model:
\begin{equation}
\mB_i(q) \ddot{q} + \mC_i(q,\dot{q})\dot{q} + \mG_{iT}(q) = {\bf \tau} + \mJ^T_{iu}F_s + \mJ^T_{ia}F_a - \mB_{iT}{q}\dot{V}_T - \mC_{iT}(q,\dot{q}){\bf V_T},
\end{equation}
where, ${\bf V_T}$ is the trunk velocity, $\mG_{iT}$ the gravitational effect of the trunk on the limb, $\mB_{iT}$ the inertia of the trunk acting on the limb, and $\mC_{iT}$ the trunk Coriolis and centrifugal term.  Note that ${\bf V_T}$ and $\dot{{\bf V}}_{\bf T}$ are directly measured through the accelerometer and gyroscope in the trunk.

The feedback linearizing controller in \cite{} operates under two fundamental assumptions and only appears to relate directly to the case of a single manipulator arm.  The first assumption of the controller is that the mass and inertial properties of an object being manipulated are estimated online.  The second is that the forces measured between the operator and suit can be modeled as
\begin{equation}
\vF_s = \mM_p \dot{\vV}_e + \vG_p,
\label{eq:forceEq}
\end{equation}
where $\mM_p$ is an effective mass of the end effector, $\dot{\vV}_e$ is the acceleration of the end effector, and $\mG_p$ is an effective gravitational term.  Note that the gravitational term is zero if the arm is not holding an external load.  The importance of these two assumptions will be more clear after the analytical forms of the feedback linearizing controllers are introduced.

The dynamics of the arm when not manipulating a load are
\begin{equation}
\mB(\vq) \ddot{\vq} + \mC(\vq,\dot{\vq})\dot{\vq} + \mG(\vq) = {\bf \tau}. \notag
\end{equation} 
The feedback linearizing controller for this case has the form
\begin{equation}
{\bf \tau} = \mB(\vq)(\ddot{\vq}_d + K_v \dot{\tilde{\vq}} + K_p \tilde{\vq}) + \mC(\vq,\dot{\vq})\dot{\vq} + \mG(\vq), \notag
\end{equation}
where $\tilde{\vq} = \vq_d - \vq$.  The details are unclear in \cite{}, but it appears that the authors either propose using predefined trajectories to define $\ddot{\vq}_d$, $\dot{\vq}_d$, and $\vq_d$, or use the assumption in \eqref{eq:forceEq} to determine these values.  
%\begin{equation}
%\vF_s = \mM_p \dot{V}_e
%\label{eq:assump1}
%\end{equation}
Equation \ref{eq:forceEq} relates the measured force to the derivative of the end effector velocity $\dot{\vV}_e$.  This end effector velocity can then be related to the desired joint velocities using the analytical Jacobian transformation \[ \ddot{\vq}_d = \mJ(\vq)^{-1}(\dot{V}_e - \dot{J}(\vq,\dot{\vq})\dot{\vq}).\]  The values for $\dot{\vq}_d$ and $\vq_d$ can then be found through integration.  It is not clear the how authors explicitly decouple the human dynamics from the robot in this case.

In the case where the exoskeleton arm is manipulating a load, the feedback linearizing controller has the augmented form
\begin{equation}
{\bf \tau} = \mB(\vq)(\ddot{\vq}_d + K_v \dot{\tilde{\vq}} + K_p \tilde{\vq}) + \mC(\vq,\dot{\vq})\dot{\vq} + \mG(\vq) -\mJ^T(\vq) \vF_s -\mJ^T(\vq)\vF_a. \notag
\end{equation}
The main difference in this case is that the controller here explicitly compensates for the measured human-robot interaction force $\vF_s$, and the controller includes the un-sensed forces $\vF_a$ which arise from the object being manipulated.  The object interaction forces thus need to be estimated.  The ability to estimate these forces is the first fundamental assumption on which the Body Extender's control scheme relies.  Because the point at which the manipulator arm and object are connected is assumed to be well known, and the arm is instrumented with inertial sensors, estimating the interaction forces of the manipulated object is be assumed to be equivalent to estimating its inertial properties.  As outlined in \cite{}, this is accomplished using a standard regressor which uses the input and output of the feedback linearizing controller as the basis for the regression.  The details of this process are left to \cite{}.

@@

The Body Extender suit is one of the only full-body exoskeletons available today, and as such is a somewhat remarkable system. In terms of how the Body Extender system is designed and how it is controlled, the suit may be limited. Clearly, for the task of handling heavy objects a fully actuated suit may be the best option, but, generally speaking, this seems almost like an over engineered solution.  The Body Extender is a heavy, overly clunky suit.  This suit may be alright for quasi-static operations, but it is unlikely that it will be useful for more than pick and place type operations.  Again, seeing as this is basically what it was designed for, that seems alright.

@@

In terms of the control of the overall suit, it was difficult to find any control papers about the Body Extender that covered anything other than upper-body single-arm control of a manipulated object.  The strategy of decoupling the individual limbs from each other, seems somewhat feasible, but is not the best way to approach coordinated limb control.  It was difficult to find literature documenting the specifics of lower-limb control for the Body Extender.  Presumably a similar strategy to that discussed here, with the addition of gravity compensation for the suit, is used.

### 2015-09-28 exos/exosText_ihmc.tex

% \begin{document}
\section{IHMC Mobility Assist Exoskeleton (MAE)}
Although the IHMC MAE was primarily designed as a disabled assist exoskeleton, the literature for this device does include a ``performance enhancement" control mode and as such is within the scope of this document.  The main design principle highlighted in the IHMC MAE hardware is a recurring theme throughout the exoskeleton literature, which is that directly measuring user intent can be difficult for a number of reasons.  This issue is resolved by using a novel control and feedback system, which in the IHMC MAE is primarily based on the use of a novel actuator design.

@@

 Actuator designed around data collected from clinical gait analysis.  Specifically, the actuators at the hip and knee were designed to be able to provide 40Nm of peak torque during the stance phase of the walking gat cycle.  The rotary SEA was composed of a Moog BN34-25EU-02 brushless DC motored with the 1:100 harmonic drive, an SHD-20.

The spring mechanism which sits between the gearbox and joint output in the IHMC MAE design is shown in Figure \ref{fig:IHMCSEA}.  The mechanism shown uses linear die springs and an angular encoder to measure the torque in the out of the joint.  The choice of linear die spring was based on the spring's predictable force to displacement function as well as favorable energy storage to weight characteristics.  
\begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/ihmc/seaAssm}
  \caption{}
  \vspace{-0.2in}
 \label{fig:IHMCSEA}   
 \end{figure}

Once assembled, the overall actuator had a  {\bf torque limit of 80 Nm}, with a {\bf velocity limit of 6.8 rad/s}.  As noted, the {\bf bandwidth of the actuator} is its main drawback, which limits the torque control to approximately {\bf 10 Hz at amplitudes greater than 15 Nm} and can track up to {\bf 30Hz for lower magnitude forces}.

@@

 The IHMC MAE is a lower extremity exoskeleton with a total of ten degrees of freedom (DOF), with five each corresponding to the two legs of the system.  Of the ten DOFs only six total are actively actuated, the other four are passive.  Each of the powered joints is powered using the SEA joint discussed in the previous subsection.  A labeled diagram of the IMHC system is shown in Figure \ref{fig:IHMCSYS}.

 The hip ad/abduction as well as flexion/extension DOFs are actively actuated.  The hip yaw DOF consists of a curved spring-loaded roller bearing whose center of rotation is approximately at the user's hip joint.

 The knee flexion/extension joint is connected to the hip flexion/extension joint using telescoping tubes and is also actuated.

 The final DOF in the IHMC design is the dorsal flexion ankle joint.  This joint is spring loaded such that it provides unidirectional torque to provide toe clearance during walking.

 In its current configuration, the IHMC exoskeleton in controlled by an off-board computer and is powered using a tether.

 The suit {\bf includes position and force sensors in the actuators in addition to two foot switches per foot}.  The foot switches are used to detect whether the system in in single or double support phase, as well as providing information on how the overall system load is approximately distributed on the stance legs.

@@

 The noted control approaches of the IHMC MAE are both relatively straight forward implementations of low-level torque based control strategies.  The first control scheme takes full advantage of the SEA actuators ability to directly measure output torque and subsequently to perform closed-loop torque control, which is pictured in \ref{fig:IHMCTOR}.     
  \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/ihmc/torqueCon}
  \caption{}
  \vspace{-0.2in}
 \label{fig:IHMCTOR}   
 \end{figure}
 The pictured control scheme takes in a desired torque, compares the desired value to the output torque measured at each individual actuator, then uses this value as well as its derivative to compute an actuator command signal, which in this case is a current.  The largest question in this controller, as presented in the literature, is how to define the desired torque term $\tau_\text{des}$.

 The second low-level control approach presented in the IHMC MAE literature is a combination of a position-based controller with an inner-loop torque controller.  This control scheme is presented in Figure \ref{fig:IHMCPOS}.
  \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/ihmc/positionCon}
  \caption{}
  \vspace{-0.2in}
 \label{fig:IHMCPOS}   
 \end{figure}
 The controller in Figure \ref{fig:IHMCPOS} shows an outer-loop position-based PD controller which takes in desired joint trajectories and compares them to measured joint positions.  The values are used to generate desired torque values which are fed into an inner closed-loop torque control loop.  This control scheme is a good low-level choice if high-impedance very accurate motions of the suit are desired.  Similarly to the torque-only control loop above, the main question in this control scheme is how the desired joint trajectories are defined.

 One possibility for defining either the desired torques in the first control scheme or the desired joint trajectories in the second is to use clinical gait analysis data.  In this approach the controllers highlighted in this subsection would act to play back nominal joint trajectory data.  The foot sensors in the system would be used to detect phase transitions and to dynamically adjust the prerecorded data to the user in realtime.

@@

 The actuator design and use of series-elastic actuation is in general a good idea for many applications, but may be a problem for the highly dynamic maneuvers required by operators in the TALOS project.  In a large sense, bandwidth is everything in these suits due to the fact that behavior prediction is very complicated.  Simply put, 10Hz control bandwidth at the actuator may be too slow to safely react in dynamically changing environments.

 The other issue with the IHMC work is that their control approach only addresses the lowest-level of control, and does not in general specify exactly how to generalize the process of of generating reference trajectories, whether they be torque or joint trajectories.  That said, other researchers do explicitly provided general methods for generating desired trajectories in what amounts to a mid-level of a hierarchical control scheme.  Thus combining other approaches with the low-level highly accurate torque control provided by the novel SEAs presented in this section could offer a potentially good solution at the actuator level.

>>>>

% \begin{document}
\section{IHMC Mobility Assist Exoskeleton (MAE)}
Although the IHMC MAE was primarily designed as a disabled assist exoskeleton, the device is also provides a ``performance enhancement" control mode, which is within the scope of this document.  The main design principle highlighted in the IHMC MAE hardware (a recurring theme throughout exoskeleton literature) is that directly measuring user intent can be difficult for a number of reasons.  The IHMC designers resolve the issue using a new control and feedback system, which for the IHMC MAE, is primarily based on a unique actuation scheme.

@@

The IHMC actuators are designed based on data collected from clinical gait analysis.  Specifically, the actuators at the hip and knee were designed to be able to provide 40Nm of peak torque during the stance phase of the walking gait cycle.  The rotary SEA is composed of a Moog BN34-25EU-02 brushless DC motor with a 1:100 harmonic drive (SHD-20).

The spring mechanism, which sits between the gearbox and joint output, is shown in Figure \ref{fig:IHMCSEA}.  The mechanism uses linear die springs and an angular encoder to measure the joint torque output.  The choice of linear die spring was based on the spring's predictable force to displacement function as well as favorable energy storage to weight characteristics.  
\begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/ihmc/seaAssm}
  \caption{}
  \vspace{-0.2in}
 \label{fig:IHMCSEA}   
 \end{figure}

Fully assembled, the actuator has a  {\bf torque limit of 80 Nm}, with a {\bf velocity limit of 6.8 rad/s}.  As noted, the {\bf bandwidth of the actuator} is its primary drawback, which limits the torque control to approximately {\bf 10 Hz at amplitudes greater than 15 Nm} and up to {\bf 30Hz for lower magnitude forces}.

@@

 The IHMC MAE lower extremity exoskeleton includes a total of ten degrees of freedom (DOF), six of which are active and four are passive.  Five DOF correspond to the two legs of the system.  Each of the powered joints is powered using the SEA joint discussed in the previous subsection.  A labeled diagram of the IMHC system is shown in Figure \ref{fig:IHMCSYS}.

 The hip ad/abduction as well as flexion/extension DOFs are actively actuated.  The hip yaw DOF consists of a curved spring-loaded roller bearing whose center of rotation is approximately at the user's hip joint. The knee flexion/extension joint is connected to the hip flexion/extension joint using telescoping tubes and is also actuated.  The final DOF is the dorsal flexion at the ankle joint.  This joint is spring loaded such that it provides unidirectional torque for toe clearance during walking.

 In its current configuration, the IHMC exoskeleton in controlled by an off-board computer and is powered using a tether. The suit {\bf includes position and force sensors in the actuators in addition to two foot switches per foot}.  The foot switches are used to detect whether the system is in single or double support phase.  It also provides information used to estimate the load distribution to the stance legs.

@@

 Both IHMC MAE control approaches noted are relatively straight forward implementations of low-level torque control.  The first control scheme takes full advantage of the SEA actuators' ability to directly measure output torque and subsequently to perform closed-loop torque control, (see \ref{fig:IHMCTOR}). 
  \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/ihmc/torqueCon}
  \caption{}
  \vspace{-0.2in}
 \label{fig:IHMCTOR}   
 \end{figure}
 The controller takes in a desired torque, compares the desired value to the output torque measured at each individual actuator, then uses this value as well as its derivative to compute an actuator command signal, which in this case is a current.  The largest question in this controller, as presented in literature, is how to define the desired torque term $\tau_\text{des}$.

 The second low-level control approach presented in the IHMC MAE literature is a combination of a position-based controller with an inner-loop torque controller.  This control scheme is presented in Figure \ref{fig:IHMCPOS}.
  \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/ihmc/positionCon}
  \caption{}
  \vspace{-0.2in}
 \label{fig:IHMCPOS}   
 \end{figure}
 The controller in Figure \ref{fig:IHMCPOS} shows an outer-loop position-based PD controller that takes in desired joint trajectories and compares them to measured joint positions.  These are used to generate desired torque values which are fed into an inner closed-loop torque control loop.  This control scheme is a good low-level choice if high-impedance. accurate suit motion is desired.  Similarly to the torque-only control loop above, the main question in this control scheme is how to obtain the desired joint trajectories.

 One possibility for defining either the desired torques / joint trajectories in these control schemes is to use clinical gait analysis data.  With this approach, the highlighted controllers would effectively play back nominal joint trajectory data.  The foot sensors in the system would detect phase transitions, allowing the controllers to dynamically adjust the prerecorded data to the user in realtime.

@@

 The design and use of series-elastic actuators is, in general, a good idea for many applications, but may be a problem for the highly dynamic maneuvers required by operators in the TALOS project.  In a large sense, bandwidth is everything in these suits due to the fact that behavior prediction is very complicated.  Simply put, 10Hz control bandwidth at the actuator may be too slow to safely react in dynamically changing environments.

 The other issue with the IHMC is that it only addresses low-level control.  The designers have not specified how to generalize the process of generating torque / joint reference trajectories.  That said, other researchers do explicitly provide methods for generating desired trajectories in what amounts to a mid-level of a hierarchical control scheme.  Thus combining other approaches with the low-level, highly accuracy torque control provided by the novel SEAs presented in this section could offer a potentially good solution at the actuator level.
